---
title: "Sejong_poll_en"
author: "coop711"
date: "`r Sys.Date()`"
output: html_document
---

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pander)
library(knitr)
options(width = 180)
```

## Data input

```{r, data}
load("./sejong_poll_data.RData")
```

### Data 

```{r, dataframe}
kable(sejong.poll.2[4:1])
```

### Factor Levels

```{r, labels in Korean}
sejong.poll.en <- sejong.poll.2
sejong.poll.en$vote <- factor(sejong.poll.en$vote, levels = c("Yes", "No"))
sejong.poll.en$class <- factor(sejong.poll.en$class, levels = class.labels)
sejong.poll.en$region <- factor(sejong.poll.en$region, levels = region.labels)
kable(sejong.poll.en[4:1])
# kable(sejong.poll.en[4:1], align = NULL, col.names = vnames.en[4:1])
```

### Cross Tables

```{r, cross tables}
tbl.vote <- xtabs(counts ~ vote, data = sejong.poll.en)
tbl.class <- xtabs(counts ~ class, data = sejong.poll.en)
tbl.region <- xtabs(counts ~ region, data = sejong.poll.en)
tbl.vote.class <- xtabs(counts ~ vote + class, data = sejong.poll.en)
tbl.vote.region <- xtabs(counts ~ vote + region, data = sejong.poll.en)
tbl.vote.class.region <- xtabs(counts ~ vote + class + region, data = sejong.poll.en)
ftbl.vote.class.region <- ftable(xtabs(counts ~ vote + class + region, data = sejong.poll.en))
```

### Array

We can set up the data as an array

```{r, array structure}
sejong.poll.array <- xtabs(counts ~ vote + class + region, data = sejong.poll.en)
str(sejong.poll.array)
sejong.poll.array
```

## Votes

### Total

Check the total vote with xtabs()

```{r, total yes or no}
vote.total <- xtabs(counts ~ vote, data = sejong.poll.en)
kable(t(as.matrix(vote.total)), caption = "Total")
# format(prop.table(vote.total)*100, digits = 3, nsmall = 1)
kable(t(as.matrix(format(prop.table(vote.total)*100, digits = 3, nsmall = 1))), caption = "Percentage", align = rep("r", 2))
vote.total.2 <- apply(sejong.poll.array, 1, sum)
# kable(t(as.matrix(vote.total.2)))
kable(t(as.matrix(vote.total.2)), caption = "Total")
```

### Vote by class

```{r, by class}
vote.class <- xtabs(counts ~ vote + class, data = sejong.poll.en)
kable(vote.class, caption = "Classwise")
vote.class.a <- apply(sejong.poll.array, 1:2, sum)
kable(vote.class.a, caption = "Classwise")
```

### Commons vs Bureaucrats

We need to analyse Commons separately. 

```{r, commons vs bureaus}
sejong.poll.en$class.2 <- factor(ifelse(sejong.poll.en$class == "Commons", "Commons", "Bureaus"), levels = c("Bureaus", "Commons"))
kable(sejong.poll.en[c(4, 3, 5, 2, 1)])
str(sejong.poll.en)
```

Compare the votes by `class.2`,  (Bureaucrats vs Commons)

```{r, vote by class.2}
vote.class.2 <- xtabs(counts ~ vote + class.2, data = sejong.poll.en)
kable(vote.class.2, caption = "Bureaus and Commons")
vote.class.2.a <- cbind("Bureaus" = rowSums(vote.class.a[, -5]), "Commons" =  vote.class.a[, 5])
kable(vote.class.2.a, caption = "Bureaus and Commons")
```

Add subtotals to the margins, 

```{r, subtotals}
vote.class.2.am <- addmargins(vote.class.2)
colnames(vote.class.2.am)[3] <- "계"
rownames(vote.class.2.am)[3] <- "계"
kable(vote.class.2.am, align = rep("r", 3))
```

Compute the marginal proportions. Note the use of `digits = 3` and `nsmall = 1`.

```{r, proportions}
kable(format(prop.table(vote.class.2, margin = 2)*100, digits = 3, nsmall = 1), caption = "Bureaus and Commons", align = rep("r", 2))
```

### Votes by region with respect to class.2 

Count the vote by region class.2 wise.

```{r, region w.r.t class.2}
class.2 <- sejong.poll.en$class.2
vote.region.bureaus <- xtabs(counts ~ vote + region, data = sejong.poll.en, class.2 == "Bureaus", drop = TRUE)
kable(vote.region.bureaus, caption = "Total of Bureaus")
# xtabs(counts ~ vote + region, data = sejong.poll.en[class.2 == "Bureaus", ], drop = TRUE)
vote.region.commons <- xtabs(counts ~ vote + region, data = sejong.poll.en, class.2 == "Commons", drop = TRUE)
kable(vote.region.commons, caption = "Total of Commons", align = rep("r", 9))
```

Seoul has three times more Bureaucrats than other regions, so analyse further.

```{r, Seoul}
region <- sejong.poll.en$region
vote.seoul.class <- xtabs(counts ~ vote + class, data = sejong.poll.en, region == "SL", drop = TRUE)
kable(vote.seoul.class, caption = "Seoul")
kable(format(prop.table(vote.seoul.class, margin = 2)*100, digits  = 3, nsmall = 1), caption = "SL", align = rep("r", 3))
```

Chungcheong's case.

```{r, Chungcheong}
vote.chung.class <- xtabs(counts ~ vote + class, data = sejong.poll.en, region == "CC", drop = TRUE)
kable(format(prop.table(vote.chung.class, margin = 2)*100, digits = 3, nsmall = 1), caption = "CC", align = rep("r", 3))
```

- Save the working directory image.

```{r, save}
save.image(file = "sejong_poll_data_en.RData")
```
