---
title: "Sejong_poll_kr"
author: "coop711"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
---

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(pander)
library(knitr)
options(width = 180)
```

## Data input

```{r, data}
load("./sejong_poll_data.RData")
```

### Data 

```{r, dataframe}
kable(sejong.poll[4:1])
```

### Korean Factor Levels

```{r, labels in Korean}
sejong.poll.kr <- sejong.poll
sejong.poll.kr$vote <- factor(sejong.poll.kr$vote, levels = c("yes", "no"), labels = c("찬성", "반대"))
class.labels.kr <- c("대신 등", "3품이하현직", "3품이하전직", "수령", "품관촌민") 
sejong.poll.kr$class <- factor(sejong.poll.kr$class, levels = class.levels, labels = class.labels.kr)
region.labels.kr <- c("서울", "유후사", "경기", "평안", "황해", "충청", "강원", "함길", "경상", "전라")
sejong.poll.kr$region <- factor(sejong.poll.kr$region, levels = region.levels, labels = region.labels.kr)
vnames.kr <- c("소계", "찬반", "계급", "지역")
# names(sejong.poll.kr) <- vnames.kr
kable(sejong.poll.kr[4:1])
# kable(sejong.poll.kr[4:1], align = NULL, col.names = vnames.kr[4:1])
```

### Cross Tables

```{r, cross tables}
tbl.vote <- xtabs(counts ~ vote, data = sejong.poll.kr)
tbl.class <- xtabs(counts ~ class, data = sejong.poll.kr)
tbl.region <- xtabs(counts ~ region, data = sejong.poll.kr)
tbl.vote.class <- xtabs(counts ~ vote + class, data = sejong.poll.kr)
tbl.vote.region <- xtabs(counts ~ vote + region, data = sejong.poll.kr)
tbl.vote.class.region <- xtabs(counts ~ vote + class + region, data = sejong.poll.kr)
ftbl.vote.class.region <- ftable(xtabs(counts ~ vote + class + region, data = sejong.poll.kr))
```

### Array

We can set up the data as an array

```{r, array structure}
sejong.poll.array <- xtabs(counts ~ vote + class + region, data = sejong.poll.kr)
str(sejong.poll.array)
sejong.poll.array
```

## Votes

### Total

Check the total vote with xtabs()

```{r, total yes or no}
vote.total <- xtabs(counts ~ vote, data = sejong.poll.kr)
kable(t(as.matrix(vote.total)), caption = "총계")
# format(prop.table(vote.total)*100, digits = 3, nsmall = 1)
kable(t(as.matrix(format(prop.table(vote.total)*100, digits = 3, nsmall = 1))), caption = "백분률", align = rep("r", 2))
vote.total.2 <- apply(sejong.poll.array, 1, sum)
# kable(t(as.matrix(vote.total.2)))
kable(t(as.matrix(vote.total.2)), caption = "총계")
```

### Vote by class

```{r, by class}
vote.class <- xtabs(counts ~ vote + class, data = sejong.poll.kr)
kable(vote.class, caption = "계급별")
vote.class.a <- apply(sejong.poll.array, 1:2, sum)
kable(vote.class.a, caption = "계급별")
```

### Commons vs Bureaucrats

We need to analyse Commons separately. 

```{r, commons vs bureaus}
sejong.poll.kr$class.2 <- factor(ifelse(sejong.poll.kr$class == "품관촌민", "품관촌민", "관료"), levels = c("관료", "품관촌민"))
kable(sejong.poll.kr[c(4, 3, 5, 2, 1)])
str(sejong.poll.kr)
```

Compare the votes by `class.2`,  (Bureaucrats vs Commons)

```{r, vote by class.2}
vote.class.2 <- xtabs(counts ~ vote + class.2, data = sejong.poll.kr)
kable(vote.class.2, caption = "관료와 품관촌민")
vote.class.2.a <- cbind("관료" = rowSums(vote.class.a[, -5]), "품관촌민" =  vote.class.a[, 5])
kable(vote.class.2.a, caption = "관료와 품관촌민")
```

Add subtotals to the margins, 

```{r, subtotals}
vote.class.2.am <- addmargins(vote.class.2)
colnames(vote.class.2.am)[3] <- "계"
rownames(vote.class.2.am)[3] <- "계"
kable(vote.class.2.am, align = rep("r", 3))
```

Compute the marginal proportions. Note the use of `digits = 3` and `nsmall = 1`.

```{r, proportions}
kable(format(prop.table(vote.class.2, margin = 2)*100, digits = 3, nsmall = 1), caption = "관료와 품관촌민", align = rep("r", 2))
```

### Votes by region with respect to class.2 

Count the vote by region class.2 wise.

```{r, region w.r.t class.2}
class.2 <- sejong.poll.kr$class.2
vote.region.bureaus <- xtabs(counts ~ vote + region, data = sejong.poll.kr, class.2 == "관료", drop = TRUE)
kable(vote.region.bureaus, caption = "집계(관료)")
# xtabs(counts ~ vote + region, data = sejong.poll.kr[class.2 == "Bureaus", ], drop = TRUE)
vote.region.commons <- xtabs(counts ~ vote + region, data = sejong.poll.kr, class.2 == "품관촌민", drop = TRUE)
kable(vote.region.commons, caption = "집계(품관촌민)", align = rep("r", 9))
```

Seoul has three times more Bureaucrats than other regions, so analyse further.

```{r, Seoul}
region <- sejong.poll.kr$region
vote.seoul.class <- xtabs(counts ~ vote + class, data = sejong.poll.kr, region == "서울", drop = TRUE)
kable(vote.seoul.class, caption = "서울")
kable(format(prop.table(vote.seoul.class, margin = 2)*100, digits  = 3, nsmall = 1), caption = "서울", align = rep("r", 3))
```

Chungcheong's case.

```{r, Chungcheong}
vote.chung.class <- xtabs(counts ~ vote + class, data = sejong.poll.kr, region == "충청", drop = TRUE)
kable(format(prop.table(vote.chung.class, margin = 2)*100, digits = 3, nsmall = 1), caption = "충청", align = rep("r", 3))
```

- Save the working directory image.

```{r, save}
save.image(file = "sejong_poll_data_kr.RData")
```
