---
title: "King Sejong's National Referendum on Tax Reform"
author: "coop711"
date: "`r Sys.Date()`"
output: html_document
---

## Data

### Reading Data 

Original data came from intenet version of Sejong silok, summarized by Oh, Ki-Soo.

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pander)
library(knitr)
options(width = 180)
```

```{r, reading data}
sejong.poll <- read.table("../data/sejong_poll.txt", header = TRUE, sep = "")
str(sejong.poll)
# pander(sejong.poll)
kable(sejong.poll[4:1])
```

### Factor conversion

We need vote, class, region as `factor`s. If you leave them as `chr`, it will be coerced to factor when you tabulate it according to alphabetical order, which is not what you want. So, use `factor()` to convert them. First, make a working copy vesion of `sejong.poll`

```{r, duplicate}
sejong.poll.2 <- sejong.poll
```

```{r, vote as factor}
sejong.poll.2$vote <- factor(sejong.poll.2$vote, levels = c("yes","no"), labels = c("Yes", "No"))
```

You can check that `labels = ` is not necessary if same as levels. Continue with class and region.

```{r, class as factor}
class.levels <- c("high","third.current", "third.ex", "chief", "ordinary")
class.labels <- c("High","3rd.current", "3rd.former", "Chief", "Commons")
sejong.poll.2$class <- factor(sejong.poll.2$class, levels = class.levels, labels = class.labels)
```

```{r, region as factor}
region.levels <- c("Seoul","yuhu", "gyunggi", "pyungan", "hwanghae", "chungcheong", "kangwon", "hamgil", "gyungsang", "jeolla")
# region.labels <- c("Seoul","Yuhu", "Gyunggi", "Pyungan", "Hwanghae", "Chungcheong", "Kangwon", "Hamgil", "Gyungsang", "Jeolla")
region.labels <- c("SL","YH", "GG", "PA", "HH", "CC", "KW", "HG", "GS", "JL")

sejong.poll.2$region <- factor(sejong.poll.2$region, levels = region.levels, labels = region.labels)
```

```{r, data in dataframe}
str(sejong.poll.2)
kable(sejong.poll.2[4:1])
```

### Array

We can set up the data as an array

```{r, array structure}
sejong.poll.array <- xtabs(counts ~ vote + class + region, data = sejong.poll.2)
str(sejong.poll.array)
sejong.poll.array
```

## Votes

### Total

Check the total vote with xtabs()

```{r, total yes or no}
vote.total <- xtabs(counts ~ vote, data = sejong.poll.2)
kable(t(as.matrix(vote.total)), caption = "Total")
# format(prop.table(vote.total)*100, digits = 3, nsmall = 1)
kable(t(as.matrix(format(prop.table(vote.total)*100, digits = 3, nsmall = 1))), caption = "Percentage", align = rep("r", 2))
vote.total.2 <- apply(sejong.poll.array, 1, sum)
# kable(t(as.matrix(vote.total.2)))
kable(t(as.matrix(vote.total.2)), caption = "Total")
```

### Vote by class

```{r, by class}
vote.class <- xtabs(counts ~ vote + class, data = sejong.poll.2)
kable(vote.class, caption = "By Class")
vote.class.a <- apply(sejong.poll.array, 1:2, sum)
kable(vote.class.a, caption = "By Class")
```

### Commons vs Bureaucrats

We need to analyse Commons separately. 

```{r, commons vs bureaus}
sejong.poll.2$class.2 <- factor(ifelse(sejong.poll.2$class == "Commons", "Commons", "Bureaus"), levels = c("Bureaus", "Commons"))
kable(sejong.poll.2[c(4, 3, 5, 2, 1)])
str(sejong.poll.2)
```

Compare the votes by `class.2`,  (Bureaucrats vs Commons)

```{r, vote by class.2}
vote.class.2 <- xtabs(counts ~ vote + class.2, data = sejong.poll.2)
kable(vote.class.2, caption = "By Bureaus and Commons")
vote.class.2.a <- cbind("Bureaus" = rowSums(vote.class.a[, -5]), "Commons" =  vote.class.a[, 5])
kable(vote.class.2.a, caption = "By Bureaus and Commons")
```

Add subtotals to the margins, 

```{r, subtotals}
vote.class.2.am <- addmargins(vote.class.2)
kable(vote.class.2.am)
```

Compute the marginal proportions. Note the use of `digits = 3` and `nsmall = 1`.

```{r, proportions}
kable(format(prop.table(vote.class.2, margin = 2)*100, digits = 3, nsmall = 1), caption = "Bureaus and Commons", align = rep("r", 2))
```

### Votes by region with respect to class.2 

Count the vote by region class.2 wise.

```{r, region w.r.t class.2}
class.2 <- sejong.poll.2$class.2
vote.region.bureaus <- xtabs(counts ~ vote + region, data = sejong.poll.2, class.2 == "Bureaus", drop = TRUE)
kable(vote.region.bureaus, caption = "Votes(Bureaus)")
# xtabs(counts ~ vote + region, data = sejong.poll.2[class.2 == "Bureaus", ], drop = TRUE)
vote.region.commons <- xtabs(counts ~ vote + region, data = sejong.poll.2, class.2 == "Commons", drop = TRUE)
kable(vote.region.commons, caption = "Votes(Commons)")
```

Seoul has three times more Bureaucrats than other regions, so analyse further.

```{r, Seoul}
region <- sejong.poll.2$region
vote.seoul.class <- xtabs(counts ~ vote + class, data = sejong.poll.2, region == "SL", drop = TRUE)
kable(vote.seoul.class, caption = "Seoul")
kable(format(prop.table(vote.seoul.class, margin = 2)*100, digits  = 3, nsmall = 1), caption = "SL", align = rep("r", 3))
```

Chungcheong's case.

```{r, Chungcheong}
vote.chung.class <- xtabs(counts ~ vote + class, data = sejong.poll.2, region == "CC", drop = TRUE)
kable(format(prop.table(vote.chung.class, margin = 2)*100, digits = 3, nsmall = 1), caption = "CC", align = rep("r", 3))
```

- Save the working directory image.

```{r, save}
save.image(file = "sejong_poll_data.RData")
```


